---
title: "190820 - TCGA DESeq analysis"
output: html_notebook
---

####Load packages:
```{r}
library(DESeq2)
```

####Load packages and settings for multi-core processing
```{r}
library(BiocParallel)
register(MulticoreParam(4))
```


####Session Info:
```{r}
sessionInfo()
```

####Load data:
```{r}
TCGA <- read.delim("/Desktop/170418_TCGA_htseqcount_Chr1_22")
cData_TCGA <- read.delim("/Desktop/cData_TCGA.txt")
colnames(TCGA)=rownames(cData_TCGA)
```

####Apply cutoff. Use at least 1 count in 1 or more samples:
```{r}
TCGA_wco=TCGA[(rowSums(TCGA>0)>=1),]
```

####DE analysis with the DESeq2 package (samples are treated as paired)
```{r}
TCGA.deseq <- DESeqDataSetFromMatrix(countData = TCGA_wco, colData = cData_TCGA, design = ~ ID + Stage)
TCGA.deseq <- DESeq(TCGA.deseq, betaPrior=TRUE, parallel=TRUE)
```

####Normalize counts by library size:
```{r}
TCGA.normalizedCounts <- t( t(counts(TCGA.deseq)) / sizeFactors(TCGA.deseq) )
```

####PCA on normalized counts
```{r}
TCGA.PCA<- prcomp(t(log2(TCGA.normalizedCounts+1)))
col.def<-c("blue","red")
colors <- rep(col.def, each=1)
M_TCGA <- summary(TCGA.PCA)
par(pty="s")
plot(TCGA.PCA$x[,1],TCGA.PCA$x[,2],col=colors,pch=20, cex=2, xlab=paste0("PC1 (", round(M_TCGA$importance[2,1]*100), "%)"),ylab=paste0("PC2 (", round(M_TCGA$importance[2,2]*100), "%)"))
text(TCGA.PCA$x[,1],TCGA.PCA$x[,2],labels=cData_TCGA$Stage, cex=0.5)
plot(TCGA.PCA$x[,3],TCGA.PCA$x[,4],col=colors,pch=20, cex=2, xlab=paste0("PC3 (", round(M_TCGA$importance[2,3]*100), "%)"),ylab=paste0("PC4 (", round(M_TCGA$importance[2,4]*100), "%)"))
text(TCGA.PCA$x[,3],TCGA.PCA$x[,4],labels=cData_TCGA$Stage, cex=0.5)
```

####Write the DE analysis results and normalized counts to a table:
```{r}
res.TCGA_wco <- results(TCGA.deseq, alpha=0.05, lfcThreshold=0, independentFiltering = TRUE, contrast = list("StagePrimary_Tumor","StageSolid_Tissue_Normal"))
write.table(res.TCGA_wco, "~/Desktop/170418_deseq_TCGA_wco_LFC0", sep="\t")
write.table(TCGA.normalizedCounts, "~/Desktop/170418_libr_size_norm_counts_TCGA", sep="\t")
```

####Volcano plot of DE analysis results
```{r}
par(pty="s")
with(res.TCGA_wco, plot(log2FoldChange, -log10(padj), pch=20))
with(subset(res.TCGA_wco, abs(log2FoldChange)>log2(2) & padj<0.05), points(log2FoldChange, -log10(padj), pch=20, col="red"))
```

####Volcano plot of DE analysis results with RBPs highlighted
```{r}
RBPs <- read.delim("~/Desktop/190823_170214_hRBP_final_list.txt")
par(pty="s")
with(res.TCGA_wco, plot(log2FoldChange, -log10(padj), pch=20))
with(subset(res.TCGA_wco, abs(log2FoldChange)>log2(2) & padj<0.05 & row.names(res.TCGA_wco)%in%RBPs[,1]), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=1))
with(subset(res.TCGA_wco, abs(log2FoldChange)<log2(2) & padj>0.05 & row.names(res.TCGA_wco)%in%RBPs[,1]), points(log2FoldChange, -log10(padj), pch=20, col="blue", cex=1))
```

####Hierarchical clustering of normalized counts
```{r}
TCGA.normalizedCounts_new_head=TCGA.normalizedCounts
colnames(TCGA.normalizedCounts_new_head)=paste0(cData_TCGA$Stage, cData_TCGA$ID)
sampleTree_TCGA_new_head <- hclust(dist(t(TCGA.normalizedCounts_new_head)), method = "average")
colLab <- function(n) { 
if (is.leaf(n)) {
a <- attributes(n)
labCol <- ifelse(grepl("Primary_Tumor",a$label),"blue","red")
attr(n, "nodePar") <- c(a$nodePar, lab.col = labCol)
}
n
}
par(mar=c(15,3,2,2))
plot(dendrapply(as.dendrogram(sampleTree_TCGA_new_head),colLab))
```
