---
title: "200701 RBP-lncRNA GO and KEGG analysis"
output: html_notebook
---

####Install packages:
```{r}
#Install packages:
#source("https://bioconductor.org/biocLite.R")
#biocLite("clusterProfiler")
#biocLite("Rgraphviz")
#biocLite("DOSE")
```

####Install specific package versions (used for initial analysis):
```{r}
#used IRanges_2.12.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/src/contrib/IRanges_2.12.0.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used S4Vectors_0.16.0.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/src/contrib/S4Vectors_0.16.0.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used BiocGenerics_0.24.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/bin/macosx/el-capitan/contrib/3.4/BiocGenerics_0.24.0.tgz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#Install old version of org.Hs.eg.db (used 3.5.0 on 190828).
packageurl <- "http://bioconductor.riken.jp/packages/3.6/data/annotation/src/contrib/org.Hs.eg.db_3.5.0.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used clusterProfiler_3.6.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/bin/macosx/el-capitan/contrib/3.4/clusterProfiler_3.6.0.tgz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used DOSE_3.4.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/bin/macosx/el-capitan/contrib/3.4/DOSE_3.4.0.tgz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used AnnotationDbi_1.40.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/bin/macosx/el-capitan/contrib/3.4/AnnotationDbi_1.40.0.tgz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
#used Biobase_2.38.0 on 190828
packageurl <- "http://bioconductor.riken.jp/packages/3.6/bioc/src/contrib/Biobase_2.38.0.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

####Load packages:
```{r, message=FALSE}
library(clusterProfiler)
library(org.Hs.eg.db)
library(DOSE)
```

```{r}
sessionInfo()
```
R version 3.4.3 (2017-11-30)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS High Sierra 10.13.6

Matrix products: default
BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.4/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] org.Hs.eg.db_3.5.0    AnnotationDbi_1.40.0  IRanges_2.12.0        S4Vectors_0.16.0      Biobase_2.38.0        BiocGenerics_0.24.0   clusterProfiler_3.6.0 DOSE_3.4.0           

loaded via a namespace (and not attached):
 [1] qvalue_2.10.0       tidyselect_1.0.0    fgsea_1.4.1         xfun_0.12           purrr_0.3.3         reshape2_1.4.3      splines_3.4.3       colorspace_1.4-1    vctrs_0.2.2         yaml_2.2.1         
[11] blob_1.2.1          rlang_0.4.4         pillar_1.4.3        glue_1.3.1          DBI_1.1.0           BiocParallel_1.12.0 bit64_0.9-7         rvcheck_0.1.7       lifecycle_0.1.0     plyr_1.8.5         
[21] stringr_1.4.0       munsell_0.5.0       GOSemSim_2.4.1      gtable_0.3.0        memoise_1.1.0       knitr_1.28          Rcpp_1.0.3          BiocManager_1.30.10 scales_1.1.0        DO.db_2.9          
[31] bit_1.1-15.2        gridExtra_2.3       fastmatch_1.1-0     ggplot2_3.2.1       digest_0.6.23       stringi_1.4.5       dplyr_0.8.4         grid_3.4.3          tools_3.4.3         magrittr_1.5       
[41] lazyeval_0.2.2      tibble_2.1.3        RSQLite_2.2.0       tidyr_1.0.2         crayon_1.3.4        GO.db_3.5.0         pkgconfig_2.0.3     data.table_1.12.0   assertthat_0.2.1    rstudioapi_0.11    
[51] R6_2.4.1            igraph_1.2.4        compiler_3.4.3 

####Load the data:
```{r}
#Note: It's better to use ENSEMBL or ENTREZ IDs for the GO analysis. The clusterProfiler package only accepts ENTREZ IDs for KEGG pathway analysis.

setwd("~/Desktop/")
RBP1 =  read.delim("190416_RBP1_FDR005_ALL_up_down_ensembl.txt", header=F)
lncFARED1 =  read.delim("200601_190426_lncFARED1_FDR001_ALL_up_down_ensembl.txt", header=F)
RBP1_lncFARED1_OVERLAP = read.delim("200601_RBP1_lncFARED1_FDR005_FDR001_ALL_up_down_OVERLAP.txt", header=F)
```

#### Biological Process GO analysis
```{r}
RBP1_BP <- enrichGO(gene = RBP1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_BP))

lncFARED1_BP <- enrichGO(gene = lncFARED1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(lncFARED1_BP))

RBP1_lncFARED1_OVERLAP_BP <- enrichGO(gene = RBP1_lncFARED1_OVERLAP$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_lncFARED1_OVERLAP_BP))
```


```{r}
#Plot top 15 BPs as dotplots
dotplot(RBP1_BP, showCategory=15)
dotplot(lncFARED1_BP, showCategory=15)
dotplot(RBP1_lncFARED1_OVERLAP_BP, showCategory=15)
```

```{r}
#Plot top 5 BPs as a cnetplot (GO terms are in brown bubbles, and gene names are in grey bubbles)
cnetplot(RBP1_BP, 5)
cnetplot(lncFARED1_BP, 5)
cnetplot(RBP1_lncFARED1_OVERLAP_BP, 5)
```

```{r}
#write the analysis results to file:
write.table(as.matrix(summary(RBP1_BP)),file="~/Desktop/200601_Summary_RBP1_BP.txt",sep="\t")
write.table(as.matrix(summary(lncFARED1_BP)),file="~/Desktop/200601_Summary_lncFARED1_BP.txt",sep="\t")
write.table(as.matrix(summary(RBP1_lncFARED1_OVERLAP_BP)),file="~/Desktop/200601_Summary_RBP1_lncFARED1_OVERLAP_BP.txt",sep="\t")
```

#### Molecular Functions GO analysis
```{r}
RBP1_MF <- enrichGO(gene = RBP1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_MF))

lncFARED1_MF <- enrichGO(gene = lncFARED1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(lncFARED1_MF))

RBP1_lncFARED1_OVERLAP_MF <- enrichGO(gene = RBP1_lncFARED1_OVERLAP$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_lncFARED1_OVERLAP_MF))
```


```{r}
#Plot top 15 MFs as dotplots
dotplot(RBP1_MF, showCategory=15)
dotplot(lncFARED1_MF, showCategory=15)
dotplot(RBP1_lncFARED1_OVERLAP_MF, showCategory=15)
```

```{r}
#Plot top 5 MFs as a cnetplot (GO terms are in brown bubbles, and gene names are in grey bubbles)
cnetplot(RBP1_MF, 5)
cnetplot(lncFARED1_MF, 5)
cnetplot(RBP1_lncFARED1_OVERLAP_MF, 5)
```

```{r}
#write the analysis results to file:
write.table(as.matrix(summary(RBP1_MF)),file="~/Desktop/200601_Summary_RBP1_MF.txt",sep="\t")
write.table(as.matrix(summary(lncFARED1_MF)),file="~/Desktop/200601_Summary_lncFARED1_MF.txt",sep="\t")
write.table(as.matrix(summary(RBP1_lncFARED1_OVERLAP_MF)),file="~/Desktop/200601_Summary_RBP1_lncFARED1_OVERLAP_MF.txt",sep="\t")
```

#### Cellular component GO analysis
```{r}
RBP1_CC <- enrichGO(gene = RBP1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_CC))

lncFARED1_CC <- enrichGO(gene = lncFARED1$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(lncFARED1_CC))

RBP1_lncFARED1_OVERLAP_CC <- enrichGO(gene = RBP1_lncFARED1_OVERLAP$V1,
                keyType       = 'ENSEMBL',
                OrgDb         = org.Hs.eg.db,
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
                minGSSize     = 5,
                readable      = TRUE)
head(summary(RBP1_lncFARED1_OVERLAP_CC))
```

```{r}
#Plot top 15 CCs as dotplots
dotplot(RBP1_CC, showCategory=15)
dotplot(lncFARED1_CC, showCategory=15)
dotplot(RBP1_lncFARED1_OVERLAP_CC, showCategory=15)
```

```{r}
#Plot top 5 CCs as a cnetplot (GO terms are in brown bubbles, and gene names are in grey bubbles)
cnetplot(RBP1_CC, 5)
cnetplot(lncFARED1_CC, 5)
cnetplot(RBP1_lncFARED1_OVERLAP_CC, 5)
```

```{r}
#write the analysis results to file:
write.table(as.matrix(summary(RBP1_CC)),file="~/Desktop/200601_Summary_RBP1_CC.txt",sep="\t")
write.table(as.matrix(summary(lncFARED1_CC)),file="~/Desktop/200601_Summary_lncFARED1_CC.txt",sep="\t")
write.table(as.matrix(summary(RBP1_lncFARED1_OVERLAP_CC)),file="~/Desktop/200601_Summary_RBP1_lncFARED1_OVERLAP_CC.txt",sep="\t")
```

#### KEGG analysis
```{r}
#First change the gene IDs from ENSEMBL to ENTREZ:
RBP1.entrez <- bitr(RBP1$V1, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Hs.eg.db)

lncFARED1.entrez <- bitr(lncFARED1$V1, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Hs.eg.db)

RBP1.lncFARED1.OVERLAP.entrez <- bitr(RBP1_lncFARED1_OVERLAP$V1, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Hs.eg.db)
```

```{r}
#Then perform the KEGG analysis
RBP1_kegg <- enrichKEGG(gene = RBP1.entrez$ENTREZID,
                 organism     = 'hsa',
                 pvalueCutoff = 0.05,
                 qvalueCutoff  = 0.05,
                 minGSSize = 3)
head(RBP1_kegg)

lncFARED1_kegg <- enrichKEGG(gene = lncFARED1.entrez$ENTREZID,
                 organism     = 'hsa',
                 pvalueCutoff = 0.05,
                 qvalueCutoff  = 0.05,
                 minGSSize = 3)
head(lncFARED1_kegg)

RBP1_lncFARED1_OVERLAP_kegg <- enrichKEGG(gene = RBP1.lncFARED1.OVERLAP.entrez$ENTREZID,
                 organism     = 'hsa',
                 pvalueCutoff = 0.05,
                 qvalueCutoff  = 0.05,
                 minGSSize = 3)
head(RBP1_lncFARED1_OVERLAP_kegg)
```


```{r}
#Plot top 15 KEGGs as dotplots
dotplot(RBP1_kegg, showCategory=15)
dotplot(lncFARED1_kegg, showCategory=15)
dotplot(RBP1_lncFARED1_OVERLAP_kegg, showCategory=15)
```

```{r}
#Plot top 5 KEGGs as a cnetplot (GO terms are in brown bubbles, and gene names are in grey bubbles)
cnetplot(RBP1_kegg, 5)
cnetplot(lncFARED1_kegg, 5)
cnetplot(RBP1_lncFARED1_OVERLAP_kegg, 5)
```

```{r}
#write the analysis results to file:
write.table(as.matrix(summary(RBP1_kegg)),file="~/Desktop/200601_Summary_RBP1_KEGG.txt",sep="\t")
write.table(as.matrix(summary(lncFARED1_kegg)),file="~/Desktop/200601_Summary_lncFARED1_KEGG.txt",sep="\t")
write.table(as.matrix(summary(RBP1_lncFARED1_OVERLAP_kegg)),file="~/Desktop/200601_Summary_RBP1_lncFARED1_OVERLAP_KEGG.txt",sep="\t")
```
