---
title: "200629 - TCGA survival analysis top 10 RBPs"
output: html_notebook
---


####Load packages:
```{r, messages=FALSE}
library(survminer)
library(dplyr)
library(survival)
library(ggrepel)
```

####Session Info
```{r}
sessionInfo()
```
R version 3.4.3 (2017-11-30)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS High Sierra 10.13.6

Matrix products: default
BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.4/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] ggrepel_0.8.1   survival_3.1-8  dplyr_0.8.4     survminer_0.4.6 ggpubr_0.2.4    magrittr_1.5    ggplot2_3.2.1   edgeR_3.20.9    limma_3.34.9   

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.3        pillar_1.4.3      compiler_3.4.3    tools_3.4.3       lifecycle_0.1.0   tibble_2.1.3      gtable_0.3.0      nlme_3.1-137      lattice_0.20-38   pkgconfig_2.0.3   rlang_0.4.4      
[12] Matrix_1.2-17     rstudioapi_0.11   yaml_2.2.1        xfun_0.12         gridExtra_2.3     withr_2.1.2       knitr_1.28        survMisc_0.5.5    generics_0.0.2    vctrs_0.2.2       locfit_1.5-9.1   
[23] grid_3.4.3        tidyselect_1.0.0  data.table_1.12.0 glue_1.3.1        KMsurv_0.1-5      R6_2.4.1          km.ci_0.5-2       purrr_0.3.3       tidyr_1.0.2       splines_3.4.3     scales_1.1.0     
[34] backports_1.1.5   assertthat_0.2.1  xtable_1.8-4      colorspace_1.4-1  ggsignif_0.6.0    lazyeval_0.2.2    munsell_0.5.0     broom_0.5.4       crayon_1.3.4      zoo_1.8-7        

####New 200420: load the original count table. Normalize. Split into lo, mid, hi in R based on tertile. Merge with metadata table

####load data
```{r}
TCGAcounts = read.delim("~/Desktop/Livernome_input/170710_all_TCGA_tumors_htseq_counts.txt", row.names = 1)

#note: since files are combined by htseq count file name, we have to load the file again, where the header is not the column names. Because R adds an X in front of numbers when used for colnames
TCGAcounts_header = read.delim("~/Desktop/Livernome_input/170710_all_TCGA_tumors_htseq_counts.txt", row.names = 1, header = FALSE)
TCGAcounts_header = TCGAcounts_header[1,]
TCGAcounts_header = t(TCGAcounts_header)
```

```{r}
TCGAmeta = read.delim("~/Desktop/Livernome_input/200420_TCGA_LIHC_metadata.txt", row.names = 6)
```

```{r}
#Merge the two tables together, in order to only analyze patients with metadata
TCGAcounts_meta = t(TCGAcounts)
rownames(TCGAcounts_meta) = TCGAcounts_header[1:425,]
TCGAcounts_meta = merge(x=TCGAcounts_meta, y=TCGAmeta, by.x=0, by.y=0, all.x=FALSE)
TCGAcounts_meta = t(TCGAcounts_meta)
```

```{r}
#Because the combined file is a character matrix we export and import again after cleaning a bit in Vim/Excel. Can also just start here with loading the 200421_TCGAcounts_withmeta_clean.txt file
write.table(as.matrix(TCGAcounts_meta),file="~/Desktop/200421_TCGAcounts_withmeta",sep="\t")
TCGAcounts_meta = read.delim("~/Desktop/Livernome_input/200421_TCGAcounts_withmeta_clean.txt", row.names = 1, header = TRUE)
```

```{r}
#Next we library normalize the counts values
TCGAcounts_colSum=colSums(TCGAcounts_meta)
TCGAcounts_LibrNorm = sweep(TCGAcounts_meta, 2, TCGAcounts_colSum, `/`)*(mean(TCGAcounts_colSum))
#colSums(TCGAcounts_LibrNorm)
```

```{r}
#Export normalized counts
write.table(as.matrix(TCGAcounts_LibrNorm),file="~/Desktop/200504_TCGAcounts_LibrNorm_withmeta",sep="\t")
```

```{r}
#convert counts to lo, mid, hi categories by expression based on tertiles. (http://www.unige.ch/ses/sococ/cl/r/groups.e.html, https://stackoverflow.com/questions/16184947/cut-error-breaks-are-not-unique)
TCGAcounts_himilo = t(TCGAcounts_LibrNorm) #make sure that it is the rows (genes) that are split, not the samples!
for(i in 1:dim(TCGAcounts_himilo)[2]) {
  breaks = quantile(TCGAcounts_himilo[,i], c(0, 1/3, 2/3, 1))
  breaks = breaks + seq_along(breaks) * .Machine$double.eps #This is to make the breaks slightly different
  TCGAcounts_himilo[,i] = cut(TCGAcounts_himilo[,i], breaks, labels = c("low", "mid", "high"), include.lowest = TRUE)
}
TCGAcounts_himilo = t(TCGAcounts_himilo)

#I'm not sure why the labels disappear when running cut on specific column. However: 1: low, 2: mid, and 3: high
#and zeroes becomes NA
```

```{r}
#convert all NA's to 1s (because they are in the low category)
TCGAcounts_himilo[is.na(TCGAcounts_himilo)] = 1
```

```{r}
#Add the header back
TCGAcounts_header = read.delim("~/Desktop/Livernome_input/200421_TCGAcounts_withmeta_clean.txt", row.names = 1, header = FALSE)
TCGAcounts_header = TCGAcounts_header[1,]
TCGAcounts_header = t(TCGAcounts_header)
```

####Merge again with metadata table
```{r, warning=FALSE}
TCGAcounts_himilo_meta = t(TCGAcounts_himilo)
colnames(TCGAcounts_himilo_meta) = TCGAcounts$gene_symbol #change ensembl to gene symbol. NOTE: not all genes have an official gene symbol in the htseq count file. Therefore a bunch of NA's are introduced here. Unless your gene of interest is among them, you can ignore this.
rownames(TCGAcounts_himilo_meta) = TCGAcounts_header[1:365,]
TCGAcounts_himilo_meta = merge(x=TCGAcounts_himilo_meta, y=TCGAmeta, by.x=0, by.y=0, all.x=TRUE)
```

```{r}
TCGAsurv = TCGAcounts_himilo_meta

#Setting reference group for Hazard Ratio computed using coxph function in R
TCGAsurv$tumor_stage_simple <- factor(TCGAsurv$tumor_stage_simple, levels = c("stage i","stage ii","stage iii","stage iv", "not reported"))
```

```{r}
TCGAsurv$RBP1 <- factor(TCGAsurv$RBP1, levels = c("1","3"))
TCGAsurv$RBP8 <- factor(TCGAsurv$RBP8, levels = c("1","3"))
TCGAsurv$RBP9 <- factor(TCGAsurv$RBP9, levels = c("1","3"))
TCGAsurv$RBP10 <- factor(TCGAsurv$RBP10, levels = c("1","3"))
TCGAsurv$RBP4 <- factor(TCGAsurv$RBP4, levels = c("1","3"))
TCGAsurv$RBP2 <- factor(TCGAsurv$RBP2, levels = c("1","3"))
TCGAsurv$RBP7 <- factor(TCGAsurv$RBP7, levels = c("1","3"))
TCGAsurv$RBP3 <- factor(TCGAsurv$RBP3, levels = c("1","3"))
TCGAsurv$RBP6 <- factor(TCGAsurv$RBP6, levels = c("1","3"))
TCGAsurv$RBP5 <- factor(TCGAsurv$RBP5, levels = c("1","3"))
```

####Base survival analysis of the TCGA-LIHC dataset
```{r}
#survival graph based on tumor stage
TCGA_surv_object <- Surv(time = TCGAsurv$time, event = TCGAsurv$censored)
fitTCGA1 <- survfit(TCGA_surv_object ~ tumor_stage_simple, data = TCGAsurv)
ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
```

```{r}
#barplots of tumor stage
par(mar=c(6,3,1,1))
barplot(table(TCGAsurv$tumor_stage_simple), las =2)
```

```{r}
#forest plot of tumor stage
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ tumor_stage_simple, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)
```

```{r}
#Age distribution in the TCGA-LIHC cohort
hist(TCGAsurv$age_at_diagnosis/365)
```

```{r}
#binning of age groups
x <- cut(TCGAsurv$age_at_diagnosis/365, breaks=c(0,40,60,75,110), labels=c("0-40","40-60","60-75","75+"))
TCGAsurv$age_at_diagnosis = x
```

```{r}
#survival graph based on age
fitTCGA1 <- survfit(TCGA_surv_object ~ age_at_diagnosis, data = TCGAsurv)
ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
```

```{r}
#forest plot of age
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ age_at_diagnosis, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)
```

```{r}
#survival graph based on gender
fitTCGA1 <- survfit(TCGA_surv_object ~ gender, data = TCGAsurv)
ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
```

```{r}
#barplots of gender
par(mar=c(6,3,1,1))
barplot(table(TCGAsurv$gender), las =2)
```

```{r}
#forest plot of gender
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ gender, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)
```

####Suvival analysis of top10 RBPs from the TCGA-LIHC (RBP1 example)
```{r}
#survival graph based on RBP1 expression (tertiles)
fitTCGA1 <- survfit(TCGA_surv_object ~ RBP1, data = TCGAsurv)
ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
```

```{r}
#forest plot of RBP1 expression (tertiles)
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ RBP1, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)
```

```{r}
#forest plot of RBP1 and metadata (adjusting for tumor-stage, age, and gender)
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ RBP1 + tumor_stage_simple + age_at_diagnosis + gender, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)
```

```{r}
#survival graph based on RBP1 expression (tertiles)
TCGAsurv$RBP1 <- factor(TCGAsurv$RBP1, levels = c("1","3"))
fitTCGA1 <- survfit(TCGA_surv_object ~ RBP1, data = TCGAsurv)
gg = ggsurvplot(fitTCGA1, 
           data = TCGAsurv,
           log.rank.weights = "1",
           pval = TRUE,
           pval.coord = c(10,4),
           pval.method = TRUE,
           pval.method.coord = c(10,10),
           censor.size = 3,
           fun = "pct",
           palette = c("#20CBF8","#000000"),
           legend = "bottom", 
           risk.table = FALSE, 
           xscale=365, 
           break.x.by=365, 
           xlab = "Time (years)")
gg
```

```{r}
#forest plot of RBP1 and metadata (adjusting for tumor-stage, age, and gender)
TCGA.fit.coxph <- coxph(TCGA_surv_object ~ RBP1 + tumor_stage_simple + age_at_diagnosis + gender, data = TCGAsurv)
ggforest(TCGA.fit.coxph, data = TCGAsurv)

```

####survival graph of all RBPs (for Fig. 1)
```{r}
#Load lists of RBP names to use
RBPlist = read.delim("~/Desktop/Livernome_input/170214_hRBP_list", row.names = 2)
RBPlist = t(RBPlist)
```

```{r}
RBP_up = read.delim("~/Desktop/Livernome_input/200506_common_UP_RBPs.txt", row.names = 2)
RBP_dn = read.delim("~/Desktop/Livernome_input/200506_common_DOWN_RBPs.txt", row.names = 2)

RBP_up = t(RBP_up)
RBP_dn = t(RBP_dn)
```


```{r}
#Upregulated RBPs
setwd("~/Desktop/200506_TCGAsurv_allRBPs/UP/")
genes<-colnames(RBP_up)
time<-TCGAsurv$time
event<-TCGAsurv$censored

for (i in 1:length(genes)){
  TCGAsurv[[genes[i]]] <- factor(TCGAsurv[[genes[i]]], levels = c("1","3"))
  pdf(paste("200506_UP_",genes[i], "_surv.pdf", sep = ""))
  fitTCGA1 <- survfit(as.formula(paste0("Surv(time , event)~",genes[i])), data = TCGAsurv)
  gg=ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
  print(gg, newpage = FALSE)
  dev.off()
}

```

```{r}
#Downregulated RBPs
setwd("~/Desktop/200506_TCGAsurv_allRBPs/DOWN")
genes<-colnames(RBP_dn)
time<-TCGAsurv$time
event<-TCGAsurv$censored

for (i in 1:length(genes)){
  TCGAsurv[[genes[i]]] <- factor(TCGAsurv[[genes[i]]], levels = c("1","3"))
  pdf(paste("200506_DN_",genes[i], "_surv.pdf", sep = ""))
  fitTCGA1 <- survfit(as.formula(paste0("Surv(time , event)~",genes[i])), data = TCGAsurv)
  gg=ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
  print(gg, newpage = FALSE)
  dev.off()
}

```


```{r}
#forest plot of all UP RBPs and metadata (adjusting for tumor-stage, age, and gender. see: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival6.html bottom example)
setwd("~/Desktop/200506_TCGAsurv_allRBPs/UP/forest/")
genes<-colnames(RBP_up)
time<-TCGAsurv$time
event<-TCGAsurv$censored
tumor_stage_simple = TCGAsurv$tumor_stage_simple
age_at_diagnosis = TCGAsurv$age_at_diagnosis
gender = TCGAsurv$gender

for (i in 1:length(genes)){
  pdf(paste("190702_",genes[i], "_forest_age_stage_gender_corrected.pdf", sep = ""))
  TCGA.fit.coxph <- coxph(as.formula(paste0("Surv(time , event)~",genes[i]," + tumor_stage_simple + age_at_diagnosis + gender")), data = TCGAsurv)
  gg=ggforest(TCGA.fit.coxph, data = TCGAsurv)
  print(gg, newpage = FALSE)
  dev.off()
}
```

```{r}
#forest plot of all DOWN RBPs and metadata (adjusting for tumor-stage, age, and gender. see: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival6.html bottom example)
setwd("~/Desktop/200506_TCGAsurv_allRBPs/DOWN/forest/")
genes<-colnames(RBP_dn)
time<-TCGAsurv$time
event<-TCGAsurv$censored
tumor_stage_simple = TCGAsurv$tumor_stage_simple
age_at_diagnosis = TCGAsurv$age_at_diagnosis
gender = TCGAsurv$gender

for (i in 1:length(genes)){
  pdf(paste("190702_",genes[i], "_forest_age_stage_gender_corrected.pdf", sep = ""))
  TCGA.fit.coxph <- coxph(as.formula(paste0("Surv(time , event)~",genes[i]," + tumor_stage_simple + age_at_diagnosis + gender")), data = TCGAsurv)
  gg=ggforest(TCGA.fit.coxph, data = TCGAsurv)
  print(gg, newpage = FALSE)
  dev.off()
}
```



```{r}
#UP data from forest plot of all RBPs and metadata (adjusting for tumor-stage, age, and gender. see: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival6.html bottom example)
setwd("~/Desktop/200506_TCGAsurv_allRBPs/UP/forest/text_data/")
genes<-colnames(RBP_up)
time<-TCGAsurv$time
event<-TCGAsurv$censored
tumor_stage_simple = TCGAsurv$tumor_stage_simple
age_at_diagnosis = TCGAsurv$age_at_diagnosis
gender = TCGAsurv$gender

for (i in 1:length(genes)){
  TCGA.fit.coxph <- coxph(as.formula(paste0("Surv(time , event)~",genes[i]," + tumor_stage_simple + age_at_diagnosis + gender")), data = TCGAsurv)
  gg=summary(TCGA.fit.coxph)
  pvc <- coef(summary(TCGA.fit.coxph))[,5]
  hr <- round(coef(summary(TCGA.fit.coxph))[,2],3)
  myfile <- file.path(paste0("200506_", genes[i], "_coxph_pval_age_stage_gender_corrected.txt"))
  write.csv(pvc, file=myfile, sep = "")
  myfile_hr <- file.path(paste0("200506_", genes[i], "_coxph_hr_age_stage_gender_corrected.txt"))
  write.csv(hr, file=myfile_hr, sep = "")
  }
```

Now go to the folder with the textfiles and cat *hr*.txt > 200506_UP_RBPs_coxph_hr_age_stage_gender_corrected.txt and cat *pval*.txt > 200506_UP_RBPs_coxph_pval_age_stage_gender_corrected.txt. These can be imported to Excel, and only unique from column A kept. Can use vlookup to merge the 2 files. Then import again:

```{r}
TCGAcoxph_UP_summary <- read.delim("~/Desktop/Livernome_input/200506_UP_RBPs_coxph_pval_hr_age_stage_gender_corrected.txt", row.names = 1)
```

```{r}
#DOWN data from forest plot of all RBPs and metadata (adjusting for tumor-stage, age, and gender. see: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival6.html bottom example)
setwd("~/Desktop/200506_TCGAsurv_allRBPs/DOWN/forest/text_data/")
genes<-colnames(RBP_dn)
time<-TCGAsurv$time
event<-TCGAsurv$censored
tumor_stage_simple = TCGAsurv$tumor_stage_simple
age_at_diagnosis = TCGAsurv$age_at_diagnosis
gender = TCGAsurv$gender

for (i in 1:length(genes)){
  TCGA.fit.coxph <- coxph(as.formula(paste0("Surv(time , event)~",genes[i]," + tumor_stage_simple + age_at_diagnosis + gender")), data = TCGAsurv)
  gg=summary(TCGA.fit.coxph)
  pvc <- coef(summary(TCGA.fit.coxph))[,5]
  hr <- round(coef(summary(TCGA.fit.coxph))[,2],3)
  myfile <- file.path(paste0("200506_", genes[i], "_coxph_pval_age_stage_gender_corrected.txt"))
  write.csv(pvc, file=myfile, sep = "")
  myfile_hr <- file.path(paste0("200506_", genes[i], "_coxph_hr_age_stage_gender_corrected.txt"))
  write.csv(hr, file=myfile_hr, sep = "")
  }
```

Now go to the folder with the textfiles and: cat *hr*.txt > 200506_DN_RBPs_coxph_hr_age_stage_gender_corrected.txt and cat *pval*.txt > 200506_DN_RBPs_coxph_pval_age_stage_gender_corrected.txt. These can be imported to Excel, and only unique from column A kept. Can use vlookup to merge the 2 files. Then import again:

```{r}
TCGAcoxph_DN_summary <- read.delim("~/Desktop/Livernome_input/200506_DN_RBPs_coxph_pval_hr_age_stage_gender_corrected.txt", row.names = 1)
```

####UP/DOWN RBPs volcano:
```{r}
RBP_UP_vp=TCGAcoxph_UP_summary
RBP_DN_vp=TCGAcoxph_DN_summary

RBP_UP_vp_sub = subset(RBP_UP_vp, hr>0.1 & pval<0.5)
RBP_DN_vp_sub = subset(RBP_DN_vp, hr<-0.1 & pval<0.5)

RBP_UPDN_vp_plot = ggplot(RBP_UP_vp) +
  geom_point(
      data = RBP_UP_vp,
      aes(x = log2(hr), y = -log10(pval)),
      fill = "blue",
      color = "black",
      cex = 2,
      pch = 21
    ) +
  geom_point(
      data = RBP_DN_vp,
      aes(x = log2(hr), y = -log10(pval)),
      fill = "red",
      color = "black",
      cex = 2,
      pch = 21
    ) +
  geom_text_repel(
      data = RBP_UP_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBP_UP_vp_sub)),
      size = 5,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
  geom_text_repel(
      data = RBP_DN_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBP_DN_vp_sub)),
      size = 5,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
      theme_bw(base_size = 14)
      
RBP_UPDN_vp_plot
```

```{r}
#Add normalized mean counts as size of the points:
name_conversion = read.delim("~/Desktop/200511_gencode.v27.annotation_conversion_table.txt", row.names = 1)
TCGAcounts_LibrNorm_rowmean = rowMeans(TCGAcounts_LibrNorm)
TCGAcounts_LibrNorm_geneid = merge(x=TCGAcounts_LibrNorm_rowmean, y=name_conversion, by.x=0, by.y=0, all.x=FALSE)

TCGAcoxph_DN_summary_count = merge(x=TCGAcoxph_DN_summary, y=TCGAcounts_LibrNorm_geneid, by.x=0, by.y=4, all.x=TRUE)
row.names(TCGAcoxph_DN_summary_count) = TCGAcoxph_DN_summary_count$Row.names

TCGAcoxph_UP_summary_count = merge(x=TCGAcoxph_UP_summary, y=TCGAcounts_LibrNorm_geneid, by.x=0, by.y=4, all.x=TRUE)
row.names(TCGAcoxph_UP_summary_count) = TCGAcoxph_UP_summary_count$Row.names
```

```{r}
#Scale the normalized counts from 0-1:
library(scales)
TCGAcoxph_DN_summary_count$relExp = rescale(TCGAcoxph_DN_summary_count$x, to = c(0, 1))
TCGAcoxph_UP_summary_count$relExp = rescale(TCGAcoxph_UP_summary_count$x, to = c(0, 1))
```


####UP/DOWN RBPs volcano (with size added):
```{r}
RBP_UP_vp=TCGAcoxph_UP_summary_count
RBP_DN_vp=TCGAcoxph_DN_summary_count
RBP_UPDN_vp = rbind(RBP_UP_vp, RBP_DN_vp)

RBP_UP_vp_sub = subset(RBP_UP_vp, hr>1 & pval<0.05)
RBP_DN_vp_sub = subset(RBP_DN_vp, hr<abs(1) & pval<0.05)

size_up = TCGAcoxph_UP_summary_count$relExp
size_dn = TCGAcoxph_DN_summary_count$relExp

RBP_UPDN_vp_plot = ggplot(RBP_UP_vp,
                          aes(x = log2(hr), y = -log10(pval))
                          ) +
  geom_point(
      aes(size = size_up),
      alpha = 0.8,
      fill = "darkgreen",
      color = "black",
      pch = 21
    ) +
  geom_point(
      data = RBP_DN_vp,
      aes(x = log2(hr), y = -log10(pval), size = size_dn),
      alpha = 0.8,
      fill = "grey",
      color = "black",
      pch = 21
    ) +
  geom_text_repel(
      data = RBP_UP_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBP_UP_vp_sub)),
      size = 4,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
  geom_text_repel(
      data = RBP_DN_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBP_DN_vp_sub)),
      size = 4,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
      theme_bw(base_size = 14) +
  labs(size="rel.exp.", x = "log2(hazard ratio)") +
  guides(size=guide_legend(override.aes=list(fill="white")))

RBP_UPDN_vp_plot
```

```{r}
RBP_UP_vp=TCGAcoxph_UP_summary_count
RBP_DN_vp=TCGAcoxph_DN_summary_count
RBP_UPDN_vp = rbind(RBP_UP_vp, RBP_DN_vp)

size_updn = log10(RBP_UPDN_vp$x)

ggplot(RBP_UPDN_vp,
             aes(x = log2(hr), y = -log10(pval))) + 
             geom_point(aes(size = size_updn), fill = "blue", color = "black", pch = 21) +
  geom_point(
      data = RBP_DN_vp,
      aes(x = log2(hr), y = -log10(pval), size = size_dn),
      size = size_dn,
      alpha = size_dn/5,
      fill = "red",
      color = "black",
      cex = 1,
      pch = 21
  ) +
             theme_bw(base_size = 14) + 
            labs(size="log10(count)")

```

####hazard volcanoes of top10 RBP familiess (for appendix)
```{r}
RBP_family = read.delim("~/Desktop/200629_TOP10_RBP_families.txt", row.names = 2)

RBP_family = t(RBP_family)
```

```{r}
setwd("~/Desktop/200629_TOP10_RBP_families/")
genes<-colnames(RBP_family)
time<-TCGAsurv$time
event<-TCGAsurv$censored

for (i in 1:length(genes)){
  TCGAsurv[[genes[i]]] <- factor(TCGAsurv[[genes[i]]], levels = c("1","3"))
  pdf(paste("200629_",genes[i], "_surv.pdf", sep = ""))
  fitTCGA1 <- survfit(as.formula(paste0("Surv(time , event)~",genes[i])), data = TCGAsurv)
  gg=ggsurvplot(fitTCGA1, data = TCGAsurv, pval = FALSE, legend = "bottom", risk.table = FALSE, xscale=365, break.x.by=365, xlab = "Time (years)")
  print(gg, newpage = FALSE)
  dev.off()
}

```

```{r}
#forest plot of RBP families and metadata (adjusting for tumor-stage, age, and gender. see: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival6.html bottom example)
setwd("~/Desktop/200629_TOP10_RBP_families/forest/")
genes<-colnames(RBP_family)
time<-TCGAsurv$time
event<-TCGAsurv$censored
tumor_stage_simple = TCGAsurv$tumor_stage_simple
age_at_diagnosis = TCGAsurv$age_at_diagnosis
gender = TCGAsurv$gender

for (i in 1:length(genes)){
  TCGA.fit.coxph <- coxph(as.formula(paste0("Surv(time , event)~",genes[i]," + tumor_stage_simple + age_at_diagnosis + gender")), data = TCGAsurv)
  gg=summary(TCGA.fit.coxph)
  pvc <- coef(summary(TCGA.fit.coxph))[,5]
  hr <- round(coef(summary(TCGA.fit.coxph))[,2],3)
  myfile <- file.path(paste0("200629_", genes[i], "_coxph_pval_age_stage_gender_corrected.txt"))
  write.csv(pvc, file=myfile, sep = "")
  myfile_hr <- file.path(paste0("200629_", genes[i], "_coxph_hr_age_stage_gender_corrected.txt"))
  write.csv(hr, file=myfile_hr, sep = "")
  }
```

Now go to the folder with the textfiles and cat *hr*.txt > 200629_TOP10_RBPfamilies_coxph_hr_age_stage_gender_corrected.txt and cat *pval*.txt > 200629_TOP10_RBPfamilies_coxph_pval_age_stage_gender_corrected.txt. These can be imported to Excel, and only unique from column A kept. Can use vlookup to merge the 2 files. Then import again:

```{r}
TCGAcoxph_RBPfamily_summary <- read.delim("~/Desktop/Livernome_input/200629_TOP10_RBPfamilies_coxph_pval_hr_age_stage_gender_corrected.txt", row.names = 2)
```

####TOP10 RBP families coxph volcano:
```{r}
TCGAcoxph_RBPfamily_summary_vp=TCGAcoxph_RBPfamily_summary

TCGAcoxph_RBPfamily_summary_vp_sub = subset(TCGAcoxph_RBPfamily_summary_vp, TCGAcoxph_RBPfamily_summary_vp$family=="RBP1")

TCGAcoxph_RBPfamily_summary_vp_plot = ggplot(TCGAcoxph_RBPfamily_summary_vp_sub) +
  geom_point(
      data = TCGAcoxph_RBPfamily_summary_vp_sub,
      aes(x = log2(hr), y = -log10(pval)),
      fill = "blue",
      color = "black",
      cex = 2,
      pch = 21
    ) +
  geom_text_repel(
      data = TCGAcoxph_RBPfamily_summary_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(TCGAcoxph_RBPfamily_summary_vp_sub)),
      size = 5,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
      theme_bw(base_size = 14)
      
TCGAcoxph_RBPfamily_summary_vp_plot
```

```{r}
#Add normalized mean counts as size of the points:
name_conversion = read.delim("~/Desktop/Livernome_input/200511_gencode.v27.annotation_conversion_table.txt", row.names = 1)
TCGAcounts_LibrNorm_rowmean = rowMeans(TCGAcounts_LibrNorm)
TCGAcounts_LibrNorm_geneid = merge(x=TCGAcounts_LibrNorm_rowmean, y=name_conversion, by.x=0, by.y=0, all.x=FALSE)

TCGAcoxph_RBPfamily_summary_count = merge(x=TCGAcoxph_RBPfamily_summary, y=TCGAcounts_LibrNorm_geneid, by.x=0, by.y=4, all.x=TRUE)
row.names(TCGAcoxph_RBPfamily_summary_count) = TCGAcoxph_RBPfamily_summary_count$Row.names

TCGAcoxph_RBPfamily_summary_count = TCGAcoxph_RBPfamily_summary_count[,c(2:5,7)]
```

```{r}
#Scale the normalized counts from 0-1:
library(scales)
TCGAcoxph_RBPfamily_summary_count$relExp = rescale(TCGAcoxph_RBPfamily_summary_count$x, to = c(0, 1))
```

####TOP10 RBP families coxph volcano:
```{r}
RBPfamily_vp=TCGAcoxph_RBPfamily_summary_count
RBPfamily_vp_sub = subset(RBPfamily_vp, hr>1 & pval<0.05)
size_RBPfamily = TCGAcoxph_RBPfamily_summary_count$relExp

RBPfamily_vp_plot = ggplot(RBPfamily_vp,
                          aes(x = log2(hr), y = -log10(pval))
                          ) +
  geom_point(
      aes(size = size_RBPfamily),
      alpha = 0.8,
      fill = "darkgreen",
      color = "black",
      pch = 21
    ) +
  geom_text_repel(
      data = RBPfamily_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBPfamily_vp_sub)),
      size = 4,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
      theme_bw(base_size = 14) +
  labs(size="rel.exp.", x = "log2(hazard ratio)") +
  guides(size=guide_legend(override.aes=list(fill="white")))

RBPfamily_vp_plot
```

####RBP1 family coxph volcano:
```{r}
RBPfamily_vp=TCGAcoxph_RBPfamily_summary_count
RBPfamily_vp_sub = subset(RBPfamily_vp, RBPfamily_vp$family=="RBP1")
RBPfamily_vp_sub$relExp = rescale(RBPfamily_vp_sub$x, to = c(0, 1))
size_RBPfamily = RBPfamily_vp_sub$relExp

RBPfamily_vp_plot = ggplot(RBPfamily_vp_sub,
                          aes(x = log2(hr), y = -log10(pval))
                          ) +
  geom_point(
      aes(size = size_RBPfamily),
      alpha = 0.8,
      fill = "#20CBF8",
      color = "black",
      pch = 21
    ) +
  geom_text_repel(
      data = RBPfamily_vp_sub,
      aes(x = log2(hr), y = -log10(pval), label=rownames(RBPfamily_vp_sub)),
      size = 4,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
      ) +
      theme_bw(base_size = 14) +
      xlim(-1,1.6) +
      ylim(0,5) +
      #scale_x_continuous(breaks = seq(-1, 1.5, by = 0.5)) +
  labs(size="rel.exp.", x = "log2(hazard ratio)") +
  guides(size=guide_legend(override.aes=list(fill="white")))

RBPfamily_vp_plot
```


####survival graphs RBP1 family
```{r}
setwd("~/Desktop/200630_TOP10_RBP_families/")
CCT_family = rownames(subset(RBPfamily_vp, RBPfamily_vp$family=="RBP1"))
genes<-CCT_family
time<-TCGAsurv$time
event<-TCGAsurv$censored

for (i in 1:length(genes)){
  TCGAsurv[[genes[i]]] <- factor(TCGAsurv[[genes[i]]], levels = c("1","3"))
  pdf(paste("200630_",genes[i], "_surv.pdf", sep = ""), width = 3, height = 3)
  fitTCGA1 <- survfit(as.formula(paste0("Surv(time , event)~",genes[i])), data = TCGAsurv)
  gg=ggsurvplot(fitTCGA1, 
           data = TCGAsurv,
           log.rank.weights = "1",
           pval = TRUE,
           pval.coord = c(10,4),
           pval.method = FALSE,
           pval.method.coord = c(10,10),
           censor.size = 3,
           fun = "pct",
           palette = c("#20CBF8","#000000"),
           legend = "bottom", 
           risk.table = FALSE, 
           xscale=365, 
           break.x.by=365,
           xlim=c(0,3650),
           xlab = "Time (years)")
  print(gg, newpage = FALSE)
  dev.off()
}

```
